---
title: "Using pipelines inside Shiny widgets or apps"
author: "Naren Srinivasan"
date: "11/14/2018"
runtime: shiny-prerendered
output: 
   rmarkdown::html_vignette:
     fig_width: 8
vignette: >
  %\VignetteIndexEntry{Using pipelines inside Shiny widgets or apps}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE, echo=FALSE, message=FALSE}
library(analysisPipelines)
knitr::opts_chunk$set(
    eval = FALSE
  )
```

# Pipelines in shiny apps
Pipelines can be used as part of Shiny widgets or apps. In the following example, we define a simple pipeline which generates a chart, and use that to power a shiny widget.

In this example, we emulate streaming dataset using the `shiny::reactivePoll` function and randomly sampling from an existing sample dataset in the package.

---

```{r data, echo=T, results='hide'}
  dataset <- read.csv(system.file("hotel_new.csv", package = "analysisPipelines"))
  dataset %>>% dplyr::select_if(is.factor) -> dataset 
  
  
  shinyPipeline <- AnalysisPipeline()
  shinyPipeline %>>% setLoggerDetails(target = "none")
  shinyPipeline %>>% univarCatDistPlots(uniCol = "building_type", priColor = "blue", optionalPlots = 0, storeOutput = T) ->   shinyPipeline
```
---

We then use the pipeline within the `shiny::renderPlot` function, and set the sampled data to execute the pipeline, and generate the chart. Since the data keeps changing due to our reactive poll, the expression within the `shiny::renderPlot` function keeps getting called in the reactive context.

---
```{r shiny, message=FALSE, warning=FALSE, eval = TRUE}
  sampled_data <- shiny::reactivePoll(intervalMillis = 2000,
                                        session = NULL,
                                        checkFunc = function() return(base::sample(1:100, 1)),
                                        valueFunc = function() return(dataset[sample(1:nrow(dataset), 100),]))
  shiny::renderPlot(height = 400, {
      sampled_data <- sampled_data()
      shinyPipeline %>>% setInput(input = sampled_data) -> shinyPipeline
      shinyPipeline %>>% generateOutput %>>% getOutputById("1")
    })
```
---

