---
title: "Analysis journey - Approach 1 - S4 Class + Pipeline operator - Prototype/ Example"
author: "Naren Srinivasan"
date: "6/13/2018"
output: html_document
---

```{r setup, include=FALSE}
source("./approach1_S4_Pipeline.R")
```

# Objective

* To prototype an approach using S4 classes and the pipeline operator to document an journey of analysis into a report

# Description of the approach

* A report is defined as an S4 class which contains the entire sequence of analyses performed on the dataset, along with the results of these analyses
* The output containing the name of the analysis and the result is stored as a tibble
* Analysis wrapper functions are defined over the base analysis functions, in order to work consistently with the defined S4 class
* The results are defined as promises on the input data, to enable lazy evaluation

# Example

## Iniializaing the object

* Data is to be loaded to the R session, and initialized into an object of class "AnalysisReport"

```{r}
data(iris)
analysisRecipeObject <- .AnalysisReport(input = iris)
str(analysisRecipeObject)
```


## Pipelining operations and creating reports

* Different reports with different sequences of analysis can be created using the '%>%' operator

```{r}
analysisRecipeObject %>% generateSummary(.) -> report1
str(report1)

analysisRecipeObject %>% plotBivariate(., "Sepal.Length", "Sepal.Width")-> report2

analysisRecipeObject %>% generateSummary(.) %>%
              plotBivariate(., "Sepal.Length", "Sepal.Width") -> report3
str(report3)
```

## Viewing results

```{r}
report3@output[1,2][[1]][[1]]
report3@output[2,2][[1]][[1]]

```

# Conclusion and next steps

* This approach provides a consistent way to sequence analysis to create a report
* The recipe (containing only sequence) vs. results should be decoupled. This can be done by having a separate Recipe class
* The performance impact of lazy evaluation needs be understood
* The choice of operator - %>% vs + needs to explored in terms of semantics